---
title: "ISyE 6739 Homework 1"
author: "Patrick Gardocki"
date: "2023-05-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2-19

## a)
$A' = 3,5,6,8$

## b)
$A \cap B = 1,2$

## c)
$(A \cap B ) \cup  C = 1,2,3,4,5$

## d)
$(B \cup C)' = 7,8$

## e)
$(A \cap B)' \cup C = 1,3,4,5,6,7,8$

# 2-81

## a)
$P(A) = \frac{52^8}{62^8} = 0.2448461$
```{r 2-81a}
52^8/62^8
```

## b)
$P(B) = \frac{10^8}{62^8} = 4.580011e-7$
```{r 2-81b}
10^8/62^8
```

## c)
$P(B\ge1) = 1 - P(A) = 0.7551539$
```{r 2-81c}
1 - 52^8/62^8
```

## d)
$P(B=2) =  \frac{{8\choose6} \times 10^2 \times 52^6}{62^8} = 0.2535389$
```{r 2-81d}
(choose(8,6)*10^2*52^6)/(62^8)
```


# 2-90

## a)
$P(A) = \frac{{5\choose1} \times 36^5}{36^6} = 5/36 \approx 0.138889$
```{r 2-90a}
(choose(5,1)*36^5)/36^6
```

## b)
$P(B) = \frac{{5\choose1} \times 36^5}{36^6} = 5/36 \approx 0.138889$
```{r 2-90b}
(choose(5,1)*36^5)/36^6
```

## c)
$P(A \cap B) = P(A) \cdot P(B) = \frac{25}{36^2} \approx 0.019290$
```{r 2-90c}
(5/36) * (5/36)
```

## d)
$P(A \cup B) =  P(A) + P(B) - P(A \cap B) = \frac{5}{36} + \frac{5}{36} - \frac{25}{36^2} = 0.258487$
```{r 2-90d}
(5/36) + (5/36) - (25/36^2)
```


# 2-141

## a)
$P(A|B') = \frac{P(A\cap B')}{P(B')} = \frac{P(A)}{1-P(B)}= \frac{P(A)}{1- P(B)}\\$
$P(A) = \frac{52^8}{62^8}; P(B) = \frac{10^8}{62^8}\\ \therefore P(A|B') = \frac{\frac{52^8}{62^8}}{1-\frac{10^8}{62^8}}= 0.24484$
```{r 2-141a}
((52^8/62^8)/(1-(10^8/62^8)))
```

## b)
$P(A' \cap B) = P(A'|B)P(B) = P(B) = \frac{10^8}{62^8} = 4.580011e-7\\$
B is a subset of A' so $P(A'|B) = 1$
```{r 2-141b}
10^8/62^8
```


## c)
Let T be passwords with exactly 2 integers and I be passwords with at least 1 integer.
$\\P(T|I) = \frac{P(T \cap I)}{P(I)} = \frac{P(T)}{P(I)}$
Since $P(T)$ is a subset of $P(I)$, $\frac{P(T \cap I)}{P(I)} = \frac{P(T)}{P(I)}\\$

From 2-81 part d, $P(T) = 0.2535\\$
From 2-81 part c, $P(I) = 0.7551\\$
$\therefore P(T|I) = \frac{0.2535}{0.7551} = 0.3357$
```{r 2-141c}
0.2535/0.7551
```

# 2-179

## a)
$P(F|S) = 0.6; P(F|V) = 0.04\\$
$P(F) = P(F|S)P(S) + P(F|V)P(V) = 0.6\times 0.2 + 0.04 \times 0.8 = 0.152$
```{r 2-179a}
0.6*0.2 + 0.8*0.04
```

## b)
$P(S|F) = \frac{P(F|S)P(S)}{P(F)} = \frac{0.6\times0.2}{0.152} = 0.78947$
```{r 2-179b}
(0.6*0.2)/0.152
```

## c)
$P(V|F') = \frac{P(F'|V)P(V)}{P(F')} = \frac{0.96\times0.8}{1-0.152} = 0.90566$
```{r 2-179c}
(0.96*0.8)/(1-0.152)
```


# Q2

${13\choose1}{4\choose3}{12\choose2}{4\choose1}{4\choose1} = 54,912$

```{r}
choose(13,1) * choose(4,3)*choose(12,2)*choose(4,1)^2
```
