---
title: "ISyE 6739 Homework 2"
author: "Patrick Gardocki"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3-109
X = number of passengers that do not show up

X follows a binomial distribution 

## a)
$P(X \ge 5) = 1 - P(X\le 4) = 1 - \sum\limits_{x=0}^{4} {125 \choose x}(0.1)^x(0.9)^{125-x} = 0.9961$
```{r 3-109a}
1 - pbinom(4,125,0.1)
```

## b) 
$P(X>5) = 1 - P(X\le 5) = 1 - \sum\limits_{x=0}^{5} {125 \choose x}(0.1)^x(0.9)^{125-x} = 0.9885$
```{r 3-109b}
1 - pbinom(5,125,0.1)
```

# 3-173

## a)
ED = number of emergency department visits per day
$\lambda T = 1.8$
$P(ED>5) = 1 - \sum\limits_{x=0}^{5}\frac{e^{-1.8}(1.8)^x}{x!} = 0.0103$
```{r 3-173a}
1 - ppois(5,1.8)
```

## b)
EW = number of emergency department visits per week
$\lambda T = 1.8 \times 7 = 12.6$
$P(ED<5) = \sum\limits_{x=0}^{4}\frac{e^{-12.6}(12.6)^x}{x!} = 0.00497$
```{r 3-173b}
ppois(4,12.6)
```

## c)
ET = number of visits in T days 
$\lambda T=1.8T$
$P(ET\ge1) = 1 -P(ET=0) = 1 - \frac{e^{-1.8T}(1.8T)^0}{0!}\\$
$e^{-1.8T} = 0.01 = ln(e^{-1.8T}) = ln(0.01) \rightarrow T = 2.558\;days$
```{r 3-173c}
log(0.01)/-1.8
```

## d)
Solve for $\lambda$.
$P(ED>5)=0.1 = 1- \sum\limits_{x=0}^{5}\frac{e^{-\lambda}(\lambda)^x}{x!}\\$


# 4-4
$f(x)=\frac{2}{x^3} \: for\:x>1$

## a)
$P(X<2) = \int\limits_{1}^{2}\frac{2}{x^3}dx= \left( \frac{-1}{x^2}\right)|_{1}^{2} = 0.75$
```{r 4-4a}
integrand <- function(x) {2/x^3}
integrate(integrand,1,2)
```

## b)
$P(X>5) = \int\limits_{5}^{\infty}\frac{2}{x^3}dx= \left( \frac{-1}{x^2}\right)|_{5}^{\infty} = 0.04$
```{r 4-4b}
integrand <- function(x) {2/x^3}
integrate(integrand,5,Inf)
```

## c)
$P(4<X<8) = \int\limits_{4}^{8}\frac{2}{x^3}dx= \left( \frac{-1}{x^2}\right)|_{4}^{8} = 0.0469$
```{r 4-4c}
integrand <- function(x) {2/x^3}
integrate(integrand,4,8)
```

## d)
$P(X<4\:or\:X>8) = \int\limits_{1}^{4}\frac{2}{x^3}dx + \int\limits_{8}^{\infty}\frac{2}{x^3}dx = 0.9531$
```{r 4-4d}
integrand <- function(x) {2/x^3}
a <- integrate(integrand,1,4)
b <- integrate(integrand,8,Inf)
a[[1]] + b[[1]]

```

## e)
Find x, where $P(X<x)=0.95$.
$\\P(X<x) = \int\limits_{1}^{x}\frac{2}{x^3} dx = \left(\frac{-1}{x^2}\right)|_{1}^{x} = \left(\frac{-1}{x^2}\right)- (-1)= 0.95\\$
$\therefore x = 4.4721$

# 4-82

## a)


