---
title: "ISyE 6739 Homework 2"
author: "Patrick Gardocki"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3-109
X = number of passengers that do not show up

X follows a binomial distribution 

## a)
$P(X \ge 5) = 1 - P(X\le 4) = 1 - \sum\limits_{x=0}^{4} {125 \choose x}(0.1)^x(0.9)^{125-x} = 0.9961$
```{r 3-109a}
1 - pbinom(4,125,0.1)
```

## b) 
$P(X>5) = 1 - P(X\le 5) = 1 - \sum\limits_{x=0}^{5} {125 \choose x}(0.1)^x(0.9)^{125-x} = 0.9885$
```{r 3-109b}
1 - pbinom(5,125,0.1)
```

# 3-173

## a)
ED = number of emergency department visits per day
$\lambda T = 1.8\\$
$P(ED>5) = 1 - \sum\limits_{x=0}^{5}\frac{e^{-1.8}(1.8)^x}{x!} = 0.0103$
```{r 3-173a}
1 - ppois(5,1.8)
```

## b)
EW = number of emergency department visits per week
$\lambda T = 1.8 \times 7 = 12.6\\$
$P(ED<5) = \sum\limits_{x=0}^{4}\frac{e^{-12.6}(12.6)^x}{x!} = 0.00497$
```{r 3-173b}
ppois(4,12.6)
```

## c)
ET = number of visits in T days 
$\lambda T=1.8T\\$
$P(ET\ge1) = 1 -P(ET=0) = 1 - \frac{e^{-1.8T}(1.8T)^0}{0!}\\$
$e^{-1.8T} = 0.01 = ln(e^{-1.8T}) = ln(0.01) \rightarrow T = 2.558\;days$
```{r 3-173c}
log(0.01)/-1.8
```

## d)
Solve for $\lambda$.
$P(ED>5)=0.1 = 1- \sum\limits_{x=0}^{5}\frac{e^{-\lambda}(\lambda)^x}{x!}\\$


# 4-4
$f(x)=\frac{2}{x^3} \: for\:x>1$

## a)
$P(X<2) = \int\limits_{1}^{2}\frac{2}{x^3}dx= \left( \frac{-1}{x^2}\right)|_{1}^{2} = 0.75$
```{r 4-4a}
integrand <- function(x) {2/x^3}
integrate(integrand,1,2)
```

## b)
$P(X>5) = \int\limits_{5}^{\infty}\frac{2}{x^3}dx= \left( \frac{-1}{x^2}\right)|_{5}^{\infty} = 0.04$
```{r 4-4b}
integrand <- function(x) {2/x^3}
integrate(integrand,5,Inf)
```

## c)
$P(4<X<8) = \int\limits_{4}^{8}\frac{2}{x^3}dx= \left( \frac{-1}{x^2}\right)|_{4}^{8} = 0.0468$
```{r 4-4c}
integrand <- function(x) {2/x^3}
integrate(integrand,4,8)
```

## d)
$P(X<4\:or\:X>8) = \int\limits_{1}^{4}\frac{2}{x^3}dx + \int\limits_{8}^{\infty}\frac{2}{x^3}dx = 0.9531$
```{r 4-4d}
integrand <- function(x) {2/x^3}
a <- integrate(integrand,1,4)
b <- integrate(integrand,8,Inf)
a[[1]] + b[[1]]

```

## e)
Find x, where $P(X<x)=0.95$.
$\\P(X<x) = \int\limits_{1}^{x}\frac{2}{x^3} dx = \left(\frac{-1}{x^2}\right)|_{1}^{x} = \left(\frac{-1}{x^2}\right)- (-1)= 0.95\\$
$\therefore x = 4.4721$

# 4-82
Let W be required water in million gallons.
Let S be storage capacity of 350 million gallons.
$\mu = 310\; million;\:\sigma = 45\: million\\$
$X\sim N(310,45^2)\\$

## a)
$P(W>S)=1-P(X\le S) = 1 - \Phi(\frac{350-310}{45}) = 0.1870$
```{r 4-82a}
pnorm(310,350,45)
```

## b) 
$P(W>w) = P(Z>\frac{w-310}{45}) = 1 - \Phi(\frac{w}{45}) = 0.01\\$
$\Phi(\frac{w-310}{45}) = 0.99 \rightarrow  \frac{w-310}{45} = 2.326$ (From Cumulative Standard Normal Table)
$\\x =414.685\: million\:gallons$
```{r 4-82 b}
qnorm(0.99)*45+310
```


## c) 
$P(W<w) = P(Z<\frac{w-310}{45}) =  \Phi(\frac{w-310}{45}) = 0.05\\$
$\Phi(\frac{w-310}{45}) = 0.05 \rightarrow  \frac{w-310}{45} = -1.644$ (From Cumulative Standard Normal Table)
$\\x =235.981\: million\:gallons$
```{r 4-82 c}
qnorm(0.05)*45+310
```

## d)
Find $\mu$, $X \sim N(\mu,45^2\\$
$P(W>S)=1-P(X\le S) = 1 - \Phi(\frac{350-\mu}{45}) =0.01\\$
$\Phi(\frac{350-\mu}{45}) = 0.99 \rightarrow \mu = 350-2.33\times45=245.15\\$
Given 1.4 million people, mean daily demand per person = 175.107 gallons 



# 4-117
$\lambda = \frac{1}{E(X)} = \frac{1}{15}$
X = time until first call $\\$

## a)
$P(X>30) = \int\limits_{30}^{\infty} \frac{1}{15}e^{\frac{-x}{15}}dx = e^{-2}=0.1353$ 
```{r 4-117 a}
integrand <- function(x) {(1/15)*exp(-x/15)}
integrate(integrand,30,Inf)
```

## b)
$P(X\le10) = P(X>10)$ Since expect 0 calls in first 10 minutes
$P(X>10) = \int\limits_{10}^{\infty} \frac{1}{15}e^{\frac{-x}{15}}dx = e^{-2/3}=0.5134$
```{r 4-117 b}
integrand <- function(x) {(1/15)*exp(-x/15)}
integrate(integrand,10,Inf)
```
## c)
$P(5<X<10) = \int\limits_{5}^{10} \frac{1}{15}e^{\frac{-x}{15}}dx = e^{-1/3}-e^{-2/3}=0.2031$
```{r 4-117 c}
integrand <- function(x) {(1/15)*exp(-x/15)}
integrate(integrand,5,10)
```
## d)
$P(X<x) = 0.90 \rightarrow P(X<x) = \int\limits_{0}^{x} \frac{1}{15}e^{\frac{-x}{15}}dx= 1-e^{-x/15}=0.90\rightarrow ln(0.10)\times-15\\$
$\therefore x=34.5387 \;minutes$
```{r 4-117 d}
log(0.1)*-15
```


# 5-6

## a)
$P(X=x, Y=y) = \frac{{60\choose 4-x-y}{30\choose y}{10\choose x}}{{100\choose 4}}\:for\:X\ge0,Y\ge0,X+Y\le4$

The table below shows the PMF table for $x,y \in \left\{ 0,1,2,3,4 \right\}$
```{r 5-6 a}
df = data.frame()
library(kableExtra)
library(knitr)
for (i in 0:4){
  c0 = choose(60,4-i)*choose(30,0)*choose(10,i)/choose(100,4)
  c1 = choose(60,4-1-i)*choose(30,1)*choose(10,i)/choose(100,4)
  c2 = choose(60,4-2-i)*choose(30,2)*choose(10,i)/choose(100,4)
  c3 = choose(60,4-3-i)*choose(30,3)*choose(10,i)/choose(100,4)
  c4 = choose(60,4-4-i)*choose(30,4)*choose(10,i)/choose(100,4)
  out = c(c0,c1,c2,c3,c4)
  df = rbind(df,out)
}

colnames(df)<-c("x=0","x=1","x=2","x=3","x=4")
row.names(df) <- c("y=0","y=1","y=2","y=3","y=4")

kable(df,"latex",align="c", caption="PMF",label="5-6a")%>%
    kable_styling(latex_options = "hold_position")
```

## b)
$f_x(x)$ is the sum of the columns in the PMF table.
$\\f_x(0) = 0.2338 \:;\: f_x(1) = 0.4187 \:;\:f_x() = 0.2679 \:;\:f_x(3) = 0.0724 \:;\:f_x(4) = 0.0069$
```{r 5-6 b}
colSums(df[ ], na.rm=TRUE)

```

## c)
$E(X) = \sum\limits_{0}^{4}x_i*f(x_i) = 1.2001$

```{r 5-6 c}
0*0.2338 + 1*0.4188 + 2*0.2679 + 3*0.0725 + 4*0.0070
```

## d)

$f_{Y|3}(y) = \frac{f_{xy}(3,y)}{f_x(3)}\\$
From 5-6 part b, using when x=3, and$f_x(3) = 0.0724$,
the PMF
```{r 5-6 d}
df1 = data.frame(y=c("0","1","2","3","4"),f=c(0.857,0.143,0,0,0))
names(df1) <- c("y","$f_{Y|3}(y)$")

kable(df1,"latex",align="c", caption="PMF",label="5-6a",escape = FALSE)%>%
    kable_styling(latex_options = "hold_position")
```



$\\$
$\\$
$\\$
$\\$

## e)
$E(Y|X=3) = \sum\limits_{0}^{4}y_i*f_{Y|3}(y)_i= 0*0.857 + 1*0.143 = 0.143$

## f)
$V(Y|X=3) = \sum\limits_{0}^{4}(y_i^2*f_{Y|3}(y)_i) - E(Y|X=3)^2 = 0^2*0.857 + 1^2*0.143 - 0.143^2 = 0.123$

## g)
X and Y are not independent.
$\\f_{xy}(x,y) \ne f_x(x)\times f_y(y)\: for\; all\; x\; and\; y\\$
For example, $f_{xy}(1,1) = 0.1354\;; f_x(1)=0.4187\;;f_y(1)=0.2994\\$
$0.1354 \ne 0.4187\times 0.2994$



# Q2

## a)
The probability of a correct lassificationa follows a binomial distribution.
$P(X) = {3\choose x}\times 0.98^x\times 0.02^{3-x}$
```{r Q2 a}
df = data.frame()
library(kableExtra)
library(knitr)
c0 = dbinom(0:3,3,0.98)
for (x in 1:4){
  
  out = c(c0[x])
  df = rbind(df,out)
}

names(df) <- c("P(x)")
row.names(df) <- c("$f(0)$","$f(1)$","$f(2)$","$f(3)$")

kable(df,"latex",align="c", caption="PMF",label="Q2",escape = FALSE)%>%
    kable_styling(latex_options = "hold_position")
```

## b)
```{r Q2 b}
c0 = pbinom(0:3,3,.98)
df = data.frame()
library(kableExtra)
library(knitr)

for (x in 1:4){
  
  out = c(c0[x])
  df = rbind(df,out)
}

names(df) <- c("P(x)")
row.names(df) <- c("$F(0)$","$F(1)$","$F(2)$","$F(3)$")

kable(df,"latex",align="c", caption="CDF",label="Q2",escape = FALSE)%>%
    kable_styling(latex_options = "hold_position")
```
$F(x) =0\;for\; x\le0\\ F(x)=1\;for\;x\ge3$


## c)
$E(X) = \sum\limits_{x=0}^{3}x* f(x) = 2.94\\$

$V(X)=\sum\limits_{x=0}^{3}x^2* f(x) -E(X)^2= 0.0588$

```{r Q2 c1}

0*.000008 + 1*0.001176 + 2*0.057624 + 3*.941192
0^2*.000008 + 1^2*0.001176 + 2^2*0.057624 + 3^2*.941192 - 2.94^2

```

