---
title: "HW3"
author: "Patrick Gardocki"
date: "2023-06-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 6-30

```{r 6-30}
octane = read.table("6-30.txt",header = TRUE)

octane=octane$Rating

o = sort(octane)

quantile(o, c(0.25,0.5,0.75),type=6)

stem(octane, scale=2)
```



# 6-46 REL DENSity is 1/2??????????

```{r 6-46}
octane = read.table("6-30.txt",header = TRUE)

octane=octane$Rating

obj = hist(octane,breaks=8)
length(octane)
library(knitr)
library(kableExtra)
df = data.frame(Frequency = obj$counts)

n=length(obj$breaks)
for(i in 1:(n-2)){
  df$Class[i]=paste(obj$breaks[i],"$\\le x <$", obj$breaks[i+1])
}

df$Class[n-1] = paste(obj$breaks[n-1],"$\\le x \\le$", obj$breaks[n])

df$"Relative Frequency"= obj$density*2
for(i in 1:(n-1)){
  df$"Cumulative Frequency"[i]=paste(cumsum(obj$counts)[i])
  df$"Cumulative Relative Frequency"[i]=paste(cumsum(obj$density*2)[i])
}

kable(df[c(2,1,3,4,5)], "latex", align="c", caption="Frecquency Distribution Table", escape = FALSE)




```


# 6-81
```{r 6-81}
Expr = read.table("6-81.txt",header = TRUE)

attach(Expr)

boxplot(Expression~Group, main="Boxplot of 4 Groups",
   xlab="Group", ylab="Expression")

```
The control groups seem to have the same median value but drastically different variances. The 'high dose' group has a mush smaller variance and slightly lower median expression as well.


# 6-98
```{r 6-98}
octane = read.table("6-30.txt",header = TRUE)

octane=octane$Rating
qqnorm(octane,datex=TRUE)
qqline(octane,datex=TRUE)

```

The Normal Probability plot suggests that the data is normally distributed but seems to break down at both ends of the data range. 

# 7-12
$\mu_X = 8.2\: minutes,\: n=49,\; \sigma_X =1.5\; minutes,\: \sigma_{\bar{x}} = \frac{\sigma_X}{\sqrt{n}}=0.2143$
Under Central Limit Theorem, $\bar{X}$ is approx. normally distributed.

## (a)
$P(\bar{X}<10)=P(Z<\frac{10-\mu}{\sigma_{\bar{X}}})=P(Z<8.4) =1$

## (b)
$P(5<\bar{X}<10)=P(\frac{5-\mu}{\sigma_{\bar{X}}}<Z<\frac{10-\mu}{\sigma_{\bar{X}}})=P(Z<8.4)-P(Z<-14.932) =1-0=1$

## (c)
$P(\bar{X}<6)=P(Z<\frac{6-\mu}{\sigma_{\bar{X}}})=P(Z<-10.27) =0$


# 7-37




# 7-44

# 8-1

# 8-41

# 8-53
